\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-21 arch.tex
updated constraint table in correct location within europa.tex and corrected arch.tex typos

% {\em\tiny This section introduces shortly T-REX architecture but move
%   quickly the focus on a single reactor and gross concepts on the
%   architecture:
%   \begin{itemize}
%   \item For t-trex the whole decision and control problem is reduced
%     around the state variables (or timeline construct). 
%   \item This reduce the execution tracking problem for each reactor on a state
%     identification problem (deduce internal state from the external
%     state), and the control problem to goal posting which will trigger
%     deliberation on the owner of the corresponding timeline. 
%   \item the architecture itself see each reactor as a black box and
%     therefore each reactor can implement its own mechanism in order to
%     resolve both synchronization and deliberation. Still we provide a
%     reactor based on the europa framework that leverage the automated
%     planning capabilities in order to do model based planning and
%     execution with a rich representation of resources and time.
%   \end{itemize}}

@@

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on {\em external} state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable.% : \comment{Needed to say that
  % early on but the flow of this paragraph is not very tight}
\kcomment{Looks fine to me unless you want to clarify further.}

@@

Requests from a reactor can be {\em recalled} when it is determined
that they are not part of its projected plan anymore. For example if a
reactor needs to replan due to an unexpected event, it will recall all
requests it produced that have not been observed \kcomment{replace
  ``observed'' with ``not incorporated by the owner reactor''?}
yet. This recall is sent to the reactor that received the goal(s) to
remove it/them from its current set of objectives.

@@

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example, one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  inconsistent. Backtracking is included in the step should it be such
  a case.  The returned value by the \texttt{step} function is a
  boolean which indicates whether the reactor needs an extra step to
  deliberate or if this last step resulted in a complete plan within
  the reactor's look-ahead. The part of the plan that covers the
  reactor's {\em external} timelines is provided to be dispatched as a
  goal.
% to whichever reactor that is responsible for these. % \comment{By
    % extending I may have saud things I developed below ... will have
    % to check}

@@

Both these deliberative \kcomment{``deliberative??''; something else
  or remove altogether perhaps?} functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference for the current tick by
identifying the current value of {\em internal} state variables from
the latest updates on {\em external} state variables. This process
propagates up in the reactor dependency graph at every tick to ensure
that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process focused on the future evolution of state variables; it can
take multiple steps spanning multiple ticks for a reactor to produce a
plan that is both complete and valid. How the reactor identifies that
its plan is complete and valid is its responsibility and depends on
the way each reactor resolves its plan at each {\em step}. The general
idea is that the reactor is able to find a valid strategy in order to
satisfy as many of its {\em internal} goals as possible. For \eu, this
means that the solver is unable to identify any further flaws to be
resolved in the plan.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
stops, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps to
target the synchronization problem by ensuring that the new {\em
  external} state updates are compatible with this plan. If so,
synchronization propagates these events in the plan and can uses it as
an aid in identifying its {\em internal} timelines \kcomment{``use it
  as an aid to identify'' what though?} based on this plan.

If they are not compatible, then {\em synchronization} identifies a
plan inconsistency. This results in the retraction of decisions made
during the \kcomment{previous?} {\em deliberation} step. % in order to
% allow {\em synchronization} to identify its {\em internal} state.
Such a relaxation can be viewed as a backtracking % on the plan search
% triggered exogenously by the world evolution. It is the way our
allowing the reactor to identify that its current plan has failed to
execute triggering further deliberations steps to find an alternate
strategy.

@@

{\em External} state updates are managed by introducing these
observations in the plan structure as facts and ensuring that they
remain compatible with the current plan mainainted by the reactor. The
challenging part of synchronization is to decide the {\em internal}
state values for the current tick. In the Shopping example for example,
we have the following code snippet which helps identify when we can
\texttt{Buy} a product:

@@

In \texttt{T-REX}'s \eu reactor this decision-making requirement is a
new type of flaw which forces identification \kcomment{identification
  of what? Probably need a little more explanation here.} of internal
state of the reactor for the current tick. This flaw can be defined
as:

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of synchronization flaws in a
    reactor. The reactor receives new observations when they are
    produced by the owner(s) of its internal timelines. The line
    \kcomment{what line???} after the last token of each timeline
    represent the domain of possible values for the end of this
    token. At every tick $\tau$ the reactor needs to integrate the
    {\em External} state information it received and -- based on its
    model $\mathcal{M}$ -- resolve its {\em Internal} state that will
    then be provided by the agent to other reactors using these state
    variables.}
  \label{fig:synch:flaw}
\end{figure}

@@

As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions \kcomment{has a unit
  decision been explained in Europa?}. Since such a process is
required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be designed carefully.

@@

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch
  window} from Definition \ref{def:dispatch} which filters out at the
agent level those goals that are before the reactor execution latency
or too far into the future. The planning window inside of a reactor
refines this filtering further, as tokens can be excluded dynamically
from deliberation as their insertion in the plan and scheduling with
other tokens generated, can reduce their temporal scope outside of the
planning scope \kcomment{ this last sentence should be reworked}. This
helps the deliberation solver reduce the number of tokens to be
evaluated reducing deliberation cost and increasing performance. As
time advances, flaws that were initially beyond this temporal horizon
will become active which a reactor can evaluate. This results in an
apparent \emph{continuous planning} of the reactor. When there are no
more flaws present in the current planning scope a partial plan
solution is found. This has two effects:

>>>>

@@

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on {\em external} state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable.

@@

Requests from a reactor can be {\em recalled} when it is determined
that they are not part of its projected plan anymore. For example if a
reactor needs to replan due to an unexpected event, it will recall all
requests produced that have yet to be executed.  This recall is also
sent to the reactor(s) that received the goal(s) to remove it/them
from its current set of objectives.

@@

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example, one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  inconsistent. Backtracking is included in the step should it be such
  a case.  The returned value by the \texttt{step} function is a
  boolean which indicates whether the reactor needs an extra step to
  deliberate or if this last step resulted in a complete plan within
  the reactor's look-ahead. The part of the plan that covers the
  reactor's {\em external} timelines is provided to be dispatched as a
  goal.

@@

Both these functions have different scope and temporal constraints.
The \texttt{synchronize} call is considered as atomic and will focus
only on state inference for the current tick by identifying the
current value of {\em internal} state variables from the latest
updates on {\em external} state variables. This process propagates up
in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process focused on
the future evolution of state variables; it can take multiple steps
spanning multiple ticks for a reactor to produce a plan that is both
complete and valid. How the reactor identifies that its plan is
complete and valid is its responsibility and depends on the way each
reactor resolves its plan at each {\em step}. The general idea is that
the reactor is able to find a valid strategy in order to satisfy as
many of its {\em internal} goals as possible. For \eu, this means that
the solver is unable to identify any further flaws to be resolved in
the plan.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
stops, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps to
target the synchronization problem by ensuring that the new {\em
  external} state updates are compatible with this plan. If so,
synchronization propagates these events in the plan and can use it as
an aid in resolving its {\em current state flaw}. It does so by using
the plan information on the exepected {\em internal} state evolution
near the execution frontier.

If they are not compatible, then {\em synchronization} identifies a
plan inconsistency. This results in the retraction of decisions made
during previous {\em deliberation} steps.  Such a relaxation can be
viewed as a backtracking allowing the reactor to identify that its
current plan has failed to execute triggering further deliberations
steps to find an alternate strategy.

@@

{\em External} state updates are managed by introducing these
observations in the plan structure as facts and ensuring that they
remain compatible with the current plan maintained by the reactor. The
challenging part of synchronization is to decide the {\em internal}
state values for the current tick. In the Shopping example for example,
we have the following code snippet which helps identify when we can
\texttt{Buy} a product:

@@

In \texttt{T-REX}'s \eu reactor this decision-making requirement is a
new type of flaw which requires identification of all values of {\em
  internal} state variables for the current tick. This flaw can be
defined as:

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\columnwidth]{figs/synch-relation}
  \caption{\small Illustration of synchronization flaws in a
    reactor. The reactor receives new observations when they are
    produced by the owner(s) of its internal timelines. The open
    ``holes'' after the last token of each timeline represents the
    domain of all possible values for the end of this token. At every
    tick $\tau$ the reactor needs to integrate the {\em External}
    state information it received and -- based on its model
    $\mathcal{M}$ -- resolve its {\em Internal} state that will then
    be provided by the agent to other reactors using these state
    variables.}
  \label{fig:synch:flaw}
\end{figure}

@@

As the example above shows, synchronization involves a deliberation
process. The risk is such that the search for a solution is not a
simple sequence of decisions where the resolution can be easily
identified by the solver as any other option would lead immediately to
n plan inconsistency. Since such a process is required by the \rx at
\emph{every} tick for \emph{every} reactor the model needs to be
designed carefully.

@@

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch window} from
Definition \ref{def:dispatch} which filters out at the agent level
those goals that are before the reactor execution latency or too far
into the future.  The planning window refines this filtering further,
as tokens can be excluded dynamically from deliberation as their
insertion in the plan and scheduling with other generated tokens can
alter their temporal scope to be outside of the planning window.  This
allows the planner to focus on its planning horizon and report
decision on flaws that are beyond this scope. As, time advances, these
potential future flaws will eventually be part of the planning window
of the reactor and will be then resolved along with potentially new
goals.  This in turn, results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

### 2012-02-21 europa.tex
updated constraint table in correct location within europa.tex and corrected arch.tex typos

\comment{explain how inference is used to detect perform bounds
  propagation, domain reduction and to detect constraint violations.} 
- talk about AC3 implementation and extension mechanisms
- talk about europa's constraint library

The Temporal Network: The notion of time is central to temporal
planning. \eu uses variables to explicitly represent timepoints for
plan activities and states. Constraints among timepoints provide a
natural way to express domain axioms. For example, in order to state
that activity A must occur before activity B we can say that the end
timepoint of A is $leq$ the start timepoint of B.

@@

\item \textbf{Reservoir} resources can have both 'consume' and
  'produce' tokens. Note that these tokens do NOT have the usual
  start/end/duration token variables, since occur at a single instant
  in time. Instead, they have a single variable, 'time' used to
  represent that instant.

\item \textbf{Reusable} resources have 'uses' tokens that use a
  quantity for the duration of the token. These tokens do have the
  usual start/end/duration variables.

\item \textbf {Unary} resources are reusable resources with unit
  quantity and have 'use' tokens (again, with the usual
  start/end/duration variables).
\end{enumerate}

\comment{talk about resource envelope computation by using MaxFlow formulation.}

>>>>

In \eu,  all the Variables and Constraints from a problem domain domain description are kept inside the Constraint Reasoning Engine (CRE).
The CRE provides a framework where any number of constraint types can collaborate to perform inference and prune variable domains using Bounds Propagation and Domain Reduction techniques. \eu uses a Propagator pattern, in which a Propagator object keeps track of a set of constraint instances of the same type and is responsible for ensuring that whenever any variable changes or constraints are introduced or dropped, all related variables are updated through inference.

In a planning problem, variable or constraint changes can come
directly from an end user (by introducing new activities or goals and
relating them to the rest of the plan through constraints, by changing
the start time of an activity, etc), or can come from an automated
planner (by expanding subgoals, introducing precedence constraints to
resolve mutual exclusion conflicts, etc). When these changes happen,
each Propagator is only able to ensure consistency for the constraints
that are associated with it.  To maintain global consistency, the CRE
implements a variant of the \texttt{AC-3} algorithm \cite{mackworth77}
where it asks Propagators to perform inference until the Constraint
Network returns to an stable state.

\comment {add europa's AC-3 algorithm here}

\input{constraints-table}

A \eu user can introduce new constraint types by implementing
specialized Constraint and Propagator classes in C++, however, \eu
provides a built-in with over 25 arithmetic, temporal, set and
resource constraint types so that a large number of problem domains
can be modeled without having to introduce new constraint
types. Tables \ref{{tab:calconst, tab:setconst,
    tab:objhierarchy,tab:misc} show the constraint types.

Temporal and resource constraints are specially important to model
real-life planning domains, so it is worth going into some more detail
about how \eu's built-in support works for those constraint types.

\textit{Temporal Constraints:} \eu uses variables to explicitly represent timepoints for
plan activities and states. Constraints among timepoints provide a
natural way to express domain axioms. For example, in order to state
that activity A must occur before activity B we can say that the end
timepoint of A is $leq$ the start timepoint of B.

@@

\item \textbf{Reservoir} resources can have both \texttt{consume(resource,time,quantity)} and
  \texttt{produce(resource,time,quantity)} constraints. Production and consumption is instantaneous (it happens at a single point in time)

\item \textbf{Reusable} resources have \texttt{uses(resource,start,duration,quantity)} constraints that use a
  quantity of the resource for a specific duration, then release it.

\item \textbf {Unary} resources are Reusable resources with capacity=1, which allows \eu to perform 
some run time savings compared to the generic Reusable class.
\end{enumerate}

\eu offers configuration options so that in addition to the usual constraint violations when a resource is completely depleted, plan flaws can be generated whenever a resource level 
is outside a specific upper or lower limit, or when instantaneous production or consumption exceeds a specific threshold.

Recall that \eu supports a flexible temporal representation, where time is represented by intervals, instead of single values. In this context, determining whether a resource is the source of a plan flaw or a constraint violation in general is more difficult than just computing a profile 
of the resource and checking against some limits. 
In the context of time flexibility, there are many resource profiles that are possible, depending on when exactly each production/consumption event happens out of all the possible values that are in the domain of the variables that represent the time for the event.
\cite{mus02} proposed an algorithm where the variables that represent the time for production/consumption events are placed in a temporal network, similar to the network used by the Temporal Propagator above, then the constraints on the resource can be seen as quantities flowing through that network. Computing the maximum flow will yield the maximum resource usage, and thus a lower bound on the resource level, and computing a minimum flow will similarly yield an upper bound. Since maximum flow on a temporal network can be computed in polynomial time, \eu takes advantage of this algorithm to efficiently compute upper and lower bound envelopes on the resource level.

\comment{figure :provide snapshot of resource envelopes generated by europa}

Resource envelopes can be very useful when trying to figure how brittle a plan in the face of changing activity durations or start times. They can also be useful when searching for plans, although they are relatively expensive to compute so their ability to prune the search space must be weighted against the ability to search faster \comment{reference Barreiro,Morris, Bresina paper on grounded inference}.
